---
title: "Cameroon Peace Promotion Project"
author: "Christopher D. Grady & Rebecca Chapman"
date: \today
output:
  word_document:
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Scan in Workspace/load libraries
load("cp3_report_dat.Rdata")
library(mosaic)
library(survey)
library(lmtest)
library(sandwich)
library(stats)
library(ggplot2)
library(reshape2)
```

# Overview

# Social Contact

Here is where we write things.  Higher scores --> more social contact.

```{r sc}
sc.plot
sc.gen.lm
```

# Cultural Understanding

Higher scores --> more understanding.

```{r cultural}
cult.plot
```

# Support for Women's Empowerment

Higher scores good.

```{r cultural}
wom.plot
wom.rel.lm

wom_money.plot
wom_work.plot
eduboys.plot
earlymarriage.plot
```

# Religious Tolerance

Higher scores good.

```{r tol}
reltol.plot

valid.plot
myrel_peace.plot
otherrel_peace.plot
diffrel_peace.plot
```

# Political/Civic Engagement

Does not form a reliable index.

```{r civic}
pol_trans.plot
dont_vote.plot
corruption.plot
comm_solve.plot

```

# Rando exp 

No plot.  Just didn't work.  Virtually everyone said yes in all conditions.

# Rel/Ethnic Feelings

Write stuff here.  These also really didn't work, since almost everyone says they overcome problems and don't vote based on ethnicity.

```{r rel/ethnic feelings}
vote_ethnic.plot
ethnic_divide.plot
religious_divide.plot
```

# Anti-Violence Empowerment

```{r anti-vio empow}
pos_change.plot
violence_problem.plot
violence_problem.plot
```

# Justice System

```{r justice system}
legal_recourse.plot
```

# Youth-Old

```{r youth-old}
youth.plot
old_respect.plot
youth_respect.plot
elders_understand.plot
elders_applicable.plot
youth_involved.plot
```



# Violence

Higher scores good, lower scores bad.

```{r tol}
vio.plot
defend_rel.plot
maintain.culture.plot
criminals_justice.plot
gov_change.plot
```




# Super Caveat

Mass Agreement Bias.  Almost everyone just agrees or says yes to every question.  This is a call for reverse coding and maybe also for more professional enumerators.  The data quality appears quite low.  OR the radio programs of PDEV really worked and these people are all super tolerant and stuff.  Well, they might be tolerant.  Or they might just know how to respond to surveys like this.
