---
title: "Cameroon Peace Promotion Project"
author: "Christopher D. Grady & Rebecca Chapman"
date: \today
output:
  word_document:
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Scan in Workspace/load libraries
load("cp3_report_dat.Rdata")
library(mosaic)
library(survey)
library(lmtest)
library(sandwich)
library(stats)
library(ggplot2)
library(reshape2)
```

# Overview

# Methodology

## Measurement Methodology

We often combine individual questions into one index that summarizes the results of those individual questions.  This is done for questions that are intended to measure the same concept -- for instance, acceptance of violence as a means to achieve goals.  Individual questions can be combined into an index when they have _internal cosistency_; that is, when a high score on one question correlates to a high score on every other question in the index.  High internal consistency indicates that the questions are all measuring one cohesive concept.  The internal consistency of the index is measure by an alpha value, with higher values meaning more internal consistency.  When an index has a high alpha value (0.7 and above), we will generally report only the index score to avoid redundancy.

We also report survey-corrected statistics.  In a survey, each respondent is not an independent data observation.  Respondents are embedded in a social context (their neighborhood) with other respondents.  Respondents in the same neighborhood tend to have similar attitudes, and so 100 respondents in one neighborhood tells us far less about an area than 100 respondents evenly spread throughout ten neighborhoods.  The 100 respondents from one neighborhood will more accurately reflect the attitudes of that particular neighborhood, but not the wider area we wish to describe.  

In this study, we surveyed 70 neighborhoods within 12 towns, or departments, in the north and extreme north regions of Cameroon.  The number of neighborhoods surveyed per town was proportional to population, so that larger towns like Garoua (11 neighborhoods) have more representation than smaller towns like Maga (2 neighborhoods).  Each neighborhood was meant to have 15 survey respondents, but due to enumerator errors the PSU sizes range from `min(tally(cpdf$psu))` to `max(tally(cpdf$psu))`.  We account for that by weighting responses such that respondents in PSUs with greater than 15 respondents are downweighted proportionally and respondents in PSUs with fewer than 15 respondents are upweighted proportionally.  That method ensures that each PSU has equal weight when calculating the town characteristics.

We also report the error of our survey statistics (i.e. means and medians) based on bootstrap replicates.  When we conduct a survey, we randomly select respondents from each neighborhood to answer survey questions.  We know the survey responses could look a little different if random selection had generated different survey respondents.  That difference is the error.  We simulate how the survey responses _might have_ looked if we conducted the survey again in the same neighborhood with a bootstrap procedure.  The bootstrap procedure is to create new hypothetical surveys by randomly selecting with replacement the respondents from our actual survey.  By conducting this bootstrapping procedure 1,000 times, we see what could have happened if we conducted 1,000 other surveys in the same locations.  That creates a distribution of possible survey statistics -- for example, a distribution of means.  We use that distribution to specify the standard error of the mean.

For example, imagine that the average score on the violence index in our survevy is 0.7.  We could imagine finding a mean of 0.65 or 0.75 if we talked to different people in that neighborhood.  However, we could _not_ imagine finding a mean of 0.2.  No combination of people in the neighborhoods would yield so low an average.  Bootstrapping quantifies the other means we could find and with what probability we would find them.

## Survey Sampling Methodology

This survey was conducted using a stratified two-level cluster sample design.



# Attitudes


## Social Contact

An important cause and consequence of intergroup tolerance is intergroup social contact.  Intergroup contact has been shown in numerous studies to increase intergroup tolerance, and more tolerant people are also more likely to seek out intergroup tolerance.  Here we measure social contact with three questions, combined into an index (alpha = 0.76).  Higher scores indicates more 

One our social contact Here, higher scores more social contact.

```{r sc}
sc.plot
#sc.gen.lm
#sc.all.lm<-svylm.fun(outcome='sc.index',treatment='gender+adult+region+religion+form.End.survey_language',data=svy)
```

## Cultural Understanding

Higher scores --> more understanding.

```{r cultural}
cult.plot
```

## Support for Women's Empowerment

### Index

Higher scores good.  Alpha only 0.59, not strong enough to consider all questions together.  Women score higher than men.  

```{r women}
wom.plot
```

### Women Household Finances

Women higher than men.  Other languages higher than French or Fulfulde.

```{r wom_money}
wom_money.plot
```

### Women Work for Pay

"When women work for pay, the children suffer".  Women score higher than men.  Non-French/Fulfulde score lower than French/Fulfulde.

```{r wom_work}
wom_work.plot
```

### Education of Boys and Girls

"Education More Important for a boy than a girl."  No statistical differences.

```{r eduboys}
eduboys.plot
```

### Early/Child Marriage

"Early marriage is good for girls."  No meaningful statistical differences.  Other religion significantly worse, but few people from that religion.

```{r earlymarriage}
earlymarriage.plot
```

## Religious Tolerance

Higher scores good.

```{r tol}
reltol.plot

valid.plot
myrel_peace.plot
otherrel_peace.plot
diffrel_peace.plot
```

## Political/Civic Engagement

Does not form a reliable index.

```{r civic}
pol_trans.plot
dont_vote.plot
corruption.plot
comm_solve.plot

```

## Rando exp 

No plot.  Just didn't work.  Virtually everyone said yes in all conditions.

## Rel/Ethnic Feelings

Write stuff here.  These also really didn't work, since almost everyone says they overcome problems and don't vote based on ethnicity.

```{r rel/ethnic feelings}
vote_ethnic.plot
ethnic_divide.plot
religious_divide.plot
```

## Anti-Violence Empowerment

```{r anti-vio empow}
pos_change.plot
violence_problem.plot
violence_problem.plot
```

## Justice System

```{r justice system}
legal_recourse.plot
```

## Youth-Old

```{r youth-old}
youth.plot
old_respect.plot
youth_respect.plot
elders_understand.plot
elders_applicable.plot
youth_involved.plot
```



## Violence

Higher scores good, lower scores bad.

```{r tol}
vio.plot
defend_rel.plot
maintain.culture.plot
criminals_justice.plot
gov_change.plot
```


# Media Access

## Utilities

```{r electricity}
electricity.plot
```

## Phones/Tech Access

### Mobile phones

Youth more than adults.  North more than Ex north.  Muslim more than Chrstian.  Fulfulde more than French, both _way_ more than other language.

```{r mobile}
mobile.plot
```

### Smartphones

French way more, men way more, North way more.  Christian more. Adults more.

```{r smartph}
smartphone.plot
```

### Computers

Only 53 people have access to a computer.  Males more likely than women to have access.  French speaking more likely than other languages.

```{r comp}
computer.plot
#tally(cpdf$computer) # 53 
```

### Tablets

Only 14 people have tablets, so these numbers mean little.  Men more than women.  French more than other languages.

```{r comp}
tablet.plot
#tally(cpdf$tablet)
```

## Mobile Service Providers

```{r msp}
mtn.plot
camtel.plot
orange.plot
nextel.plot
otherMSP.plot
```

## OS

```{r OS}
android.plot
blackberry.plot
ios.plot
windows.plot
```

## Internet Use

Zero people used the internet on tablets.

```{r internet}
int.plot
int_friend_rel.plot
int_cafe.plot
int_mobile.plot
int_tablet.plot
int_other.plot
```

## Media/Radio Listening

Only 9 people listened to Dabalaye.  Only 3 people listened to Chabab.  0 people listened to Dandal_Kura.

```{r radio programs}
don_derkeen.plot
dd_freq.plot
dd_op.plot
```

## Arewa24 Plots

```{r arewa}
alawar.plot
waiwaye.plot
dadin.plot
hhh.plot
kundin.plot
tauraruwa.plot
gari.plot
matasa.plot
other.plot
```


# Conclusion
 
Sum up findings here.

# Super Caveat

Mass Agreement Bias.  Almost everyone just agrees or says yes to every question.  This is a call for reverse coding and maybe also for more professional enumerators.  The data quality appears quite low.  OR the radio programs of PDEV really worked and these people are all super tolerant and stuff.  Well, they might be tolerant.  Or they might just know how to respond to surveys like this.
