---
title: "Cameroon Peace Promotion Project"
author: "Christopher D. Grady & Rebecca Chapman"
date: \today
output:
  word_document:
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Scan in Workspace/load libraries
load("cp3_report_dat.Rdata")
library(mosaic)
library(survey)
library(lmtest)
library(sandwich)
library(stats)
library(ggplot2)
library(reshape2)
```

# Overview

# Attitudes

## Social Contact

Here is where we write things.  Higher scores --> more social contact.

```{r sc}
sc.plot
sc.gen.lm
```

## Cultural Understanding

Higher scores --> more understanding.

```{r cultural}
cult.plot
```

## Support for Women's Empowerment

### Index

Higher scores good.  Alpha only 0.59, not strong enough to consider all questions together.  Women score higher than men.  

```{r women}
wom.plot
```

### Women Household Finances

Women higher than men.  Other languages higher than French or Fulfulde.

```{r wom_money}
wom_money.plot
```

### Women Work for Pay

"When women work for pay, the children suffer".  Women score higher than men.  Non-French/Fulfulde score lower than French/Fulfulde.

```{r wom_work}
wom_work.plot
```

### Education of Boys and Girls

"Education More Important for a boy than a girl."  No statistical differences.

```{r eduboys}
eduboys.plot
```

### Early/Child Marriage

"Early marriage is good for girls."  No meaningful statistical differences.  Other religion significantly worse, but few people from that religion.

```{r earlymarriage}
earlymarriage.plot
```

## Religious Tolerance

Higher scores good.

```{r tol}
reltol.plot

valid.plot
myrel_peace.plot
otherrel_peace.plot
diffrel_peace.plot
```

## Political/Civic Engagement

Does not form a reliable index.

```{r civic}
pol_trans.plot
dont_vote.plot
corruption.plot
comm_solve.plot

```

## Rando exp 

No plot.  Just didn't work.  Virtually everyone said yes in all conditions.

## Rel/Ethnic Feelings

Write stuff here.  These also really didn't work, since almost everyone says they overcome problems and don't vote based on ethnicity.

```{r rel/ethnic feelings}
vote_ethnic.plot
ethnic_divide.plot
religious_divide.plot
```

## Anti-Violence Empowerment

```{r anti-vio empow}
pos_change.plot
violence_problem.plot
violence_problem.plot
```

## Justice System

```{r justice system}
legal_recourse.plot
```

## Youth-Old

```{r youth-old}
youth.plot
old_respect.plot
youth_respect.plot
elders_understand.plot
elders_applicable.plot
youth_involved.plot
```



## Violence

Higher scores good, lower scores bad.

```{r tol}
vio.plot
defend_rel.plot
maintain.culture.plot
criminals_justice.plot
gov_change.plot
```


# Media Access

## Utilities

```{r electricity}
electricity.plot
```

## Phones/Tech Access

### Mobile phones

Youth more than adults.  North more than Ex north.  Muslim more than Chrstian.  Fulfulde more than French, both _way_ more than other language.

```{r mobile}
mobile.plot
```

### Smartphones

French way more, men way more, North way more.  Christian more. Adults more.

```{r smartph}
smartphone.plot
```

### Computers

Only 53 people have access to a computer.  Males more likely than women to have access.  French speaking more likely than other languages.

```{r comp}
computer.plot
#tally(cpdf$computer) # 53 
```

### Tablets

Only 14 people have tablets, so these numbers mean little.  Men more than women.  French more than other languages.

```{r comp}
tablet.plot
#tally(cpdf$tablet)
```

## Mobile Service Providers

```{r msp}
mtn.plot
camtel.plot
orange.plot
nextel.plot
otherMSP.plot
```

## OS

```{r OS}
android.plot
blackberry.plot
ios.plot
windows.plot
```

## Internet Use

Zero people used the internet on tablets.

```{r internet}
int.plot
int_friend_rel.plot
int_cafe.plot
int_mobile.plot
int_tablet.plot
int_other.plot
```

## Media/Radio Listening

Only 9 people listened to Dabalaye.  Only 3 people listened to Chabab.  0 people listened to Dandal_Kura.

```{r radio programs}
don_derkeen.plot
dd_freq.plot
dd_op.plot
```

## Arewa24 Plots

```{r arewa}
alawar.plot
waiwaye.plot
dadin.plot
hhh.plot
kundin.plot
tauraruwa.plot
gari.plot
matasa.plot
other.plot
```


# Conclusion
 
Sum up findings here.

# Super Caveat

Mass Agreement Bias.  Almost everyone just agrees or says yes to every question.  This is a call for reverse coding and maybe also for more professional enumerators.  The data quality appears quite low.  OR the radio programs of PDEV really worked and these people are all super tolerant and stuff.  Well, they might be tolerant.  Or they might just know how to respond to surveys like this.
