---
title: "Cameroon Peace Promotion Project"
author: "Christopher D. Grady & Rebecca Chapman"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include=T)

# Scan in Workspace/load libraries
load("cp3_report_dat.Rdata")
library(mosaic)
library(survey)
library(lmtest)
library(sandwich)
library(stats)
library(ggplot2)
library(reshape2)
```

# Overview

# Methodology

## Measurement Methodology

We often combine individual questions into one index that summarizes the results of those individual questions.  This is done for questions that are intended to measure the same concept -- for instance, acceptance of violence as a means to achieve goals.  Individual questions can be combined into an index when they have _internal cosistency_; that is, when a high score on one question correlates to a high score on every other question in the index.  High internal consistency indicates that the questions are all measuring one cohesive concept.  The internal consistency of the index is measure by an alpha value, with higher values meaning more internal consistency.  When an index has a high alpha value (0.7 and above), we will generally report only the index score to avoid redundancy.

The survey questions mainly use modified Likert scales.  On our modified Likert scales, respondents are presented with a statement and asked the extent of their agreement or disagreement with that question.  The options we use are (1) strongly agree, (2) somewhat agree, (3) somewhat disagree, and (4) strongly disagree.  We do not enumerate a "don't know" option to respondents, since research shows that certain demographic groups are systematically more likely to use don't know responses, even when they are fairly certain of the answer (Krosnick 2002).  Primarily, introverted people and people from disadvantaged groups (i.e. people with low education, women, etc...).  Offering a "don't know" response systematically decreases the representation of these people in the survey and increases the representation of people who are not afraid to guess even when they do not know the answer.  These extroverted people tend to be older men, and so enumerating a "don't know" response results in a survey that overrepresents older men.  We instruct enumerators not to enumerate a "don't know" response and only utilize "don't know" if the respondent asserts that they cannot or do not wish to answer.

We also report survey-corrected statistics.  In a survey, each respondent is not an independent data observation.  Respondents are embedded in a social context (their neighborhood) with other respondents.  Respondents in the same neighborhood tend to have similar attitudes, and so 100 respondents in one neighborhood tells us far less about an area than 100 respondents evenly spread throughout ten neighborhoods.  The 100 respondents from one neighborhood will more accurately reflect the attitudes of that particular neighborhood, but not the wider area we wish to describe.  

In this study, we surveyed 70 neighborhoods within 12 towns, or departments, in the north and extreme north regions of Cameroon.  The number of neighborhoods surveyed per town was proportional to population, so that larger towns like Garoua (11 neighborhoods) have more representation than smaller towns like Maga (2 neighborhoods).  Each neighborhood was meant to have 15 survey respondents, but due to enumerator errors the PSU sizes range from `min(tally(cpdf$psu))` to `max(tally(cpdf$psu))`.  We account for that by weighting responses such that respondents in PSUs with greater than 15 respondents are downweighted proportionally and respondents in PSUs with fewer than 15 respondents are upweighted proportionally.  That method ensures that each PSU has equal weight when calculating the town characteristics.

We also report the error of our survey statistics (i.e. means and medians) based on bootstrap replicates.  When we conduct a survey, we randomly select respondents from each neighborhood to answer survey questions.  We know the survey responses could look a little different if random selection had generated different survey respondents.  That difference is the error.  We simulate how the survey responses _might have_ looked if we conducted the survey again in the same neighborhood with a bootstrap procedure.  The bootstrap procedure is to create new hypothetical surveys by randomly selecting with replacement the respondents from our actual survey.  By conducting this bootstrapping procedure 1,000 times, we see what could have happened if we conducted 1,000 other surveys in the same locations.  That creates a distribution of possible survey statistics -- for example, a distribution of means.  We use that distribution to specify the standard error of the mean.

For example, imagine that the average score on the violence index in our survevy is 0.7.  We could imagine finding a mean of 0.65 or 0.75 if we talked to different people in that neighborhood.  However, we could _not_ imagine finding a mean of 0.2.  No combination of people in the neighborhoods would yield so low an average.  Bootstrapping quantifies the other means we could find and with what probability we would find them.

## Survey Sampling Methodology

This survey was conducted using a stratified two-level cluster sample design.  The survey was stratified by department/town; we have every department/town in our target area.  Within towns, we randomly select PSUs from a list obtained by workers on the ground.  Within PSUs, the enumerators mapped out 75 households and randomly selected 15 of them (1/5th) through a systematic sampling procedure.  And within selected households, enumerators randomly selected one respondent by assigning each household member a number and using a random number generator to select one number. 

If a household did not consent to be surveyed, enumerators proceeded to the next non-selected household.  If a respondent did not consent, the enumerator also proceeded to the next non-selected household.

In the next section, we summarize the findings of the survey, disaggregating answers by gender, age, region, religion, and the language spoken for the survey.  We display a graph for each question group, disaggregating by (1) gender, (2) age, (3) region, (4) religion^[Though we display estimates for Christians, Muslims, and people from other religions, only 11 people responded with a religion other than Christianity or Islam.  As such, these estimates are extremely imprecise.], and (5) interview language.  The bars show the mean response of respondents in each group, and the error bars show the 95% confidence interval within which the mean _could_ be.

# Attitudes


## Social Contact

An important cause and consequence of intergroup tolerance is intergroup social contact.  Intergroup contact has been shown in numerous studies to increase intergroup tolerance, and more tolerant people are also more likely to seek out intergroup contact.  Here we measure social contact with three questions, combined into an index (alpha = 0.76).  The index is scaled from 0-1 and higher scores indicates more contact with outgroups.  The specific questions are about frequency of contact with (1) ethnic outgroups, (2) religious outgroups, and (3) foreigners.

The mean score on the social context index is about 0.65, but there are some large differences The largest difference is between people in different regions, with people in the northern region reporting much more intergroup contact than people in the extreme northern region.  People who answered the survey in a language other than French or Fulfulde are also much less likely to report having intergroup contact.  The graph shows other differences, but these are small and either not statistically significant or not substantively meaningful.

```{r sc}
sc.plot
#summary(cpdf$sc.index)/max(cpdf$sc.index,na.rm=T)
#sc.gen.lm
#sc.all.lm<-svylm.fun(outcome='sc.index',treatment='gender+adult+region+religion+form.End.survey_language',data=svy)
```

## Cultural Understanding

A perceived lack of cultural understanding often feeds the us vs. them narrative that underlies intergroup conflict.  This survey measures both a respondent's feeling that living in harmony with different types of people is important in their culture, and the respondent's perception that various "others" understand and respect their culture.  Those others are (1) people from other religions, (2) people from other ethnicities, (3) people from other regions of Cameroon, (4) political leaders, and (5) youth.

The mean score for this index is almost 0.9 out of 1, and the graph shows that there are not large differences in perceptions of cultural understanding between most subgroups in the data.  Young people and Muslims both score more highly on this index, but only by about .04 points on the 0-1 scale.  Overall, all of these subgroups agree equally that it is important to live peacefully with others and that others understand and respect their culture.


```{r cultural}
cult.plot
#summary(cpdf$cult.index)/max(cpdf$cult.index,na.rm=T)
#cult.y.lm<-svylm.fun(outcome='cult.index',treatment='adult',data=svy)
#cult.r.lm<-svylm.fun(outcome='cult.index',treatment='religion',data=svy)
#cult.all.lm<-svylm.fun(outcome='cult.index',treatment='gender+adult+region+religion+form.End.survey_language',data=svy)
#ycult.all.lm<-svylm.fun(outcome='form.Attitudes.youth_proud_culture',treatment='gender+adult+region+religion+form.End.survey_language',data=svy) # youth more on youth
#ecult.all.lm<-svylm.fun(outcome='form.Attitudes.other_regions_respect',treatment='gender+adult+region+religion+form.End.survey_language',data=svy) # youth not more on other outcomes
```

## Support for Women's Empowerment

### Index

Women are often denied empowerment over their lives.  Women are often denied the education of their male counterparts, a say in decisions that affect their communities, or even a choice in their marriage partner. The women's empowerment index measures the extent to which respondents believe that women should have be empowered to the same extent as men.  It contains four questions: (1) A woman should have a say in how her household spends money; (2) When a mother works for pay the children suffer; (3) An education is more important for a boy than a girl; and (4) It is not in the best interest of a girl to be married before she is 16 years old.  All questions are coded so that higher scores indicate more support for women's empowerment.

The mean score for the index is about [number] on a scale from 0-1, indicating a good amount of support for women's empowerment.  Women score higher than men.  Extreme north scores higher than north.  People speaking languages other than French or Fulfulde score lower.

Unlike our previous indices, the alpha for the women's empowerment index is only 0.58.  This is not enough internal consistency of us to consider these questions as measuring the same concept.  Therefore we now turn to analyses of the individual questions in the women's empowerment index.

```{r women}
wom.plot
wom.all.lm<-svylm.fun(outcome='wom.index',treatment='gender+adult+region+religion+form.End.survey_language',data=svy)
```

### Women Household Finances

The first question in the women's empowerment sequence asks if a woman should have a say in how her household spend money.  The average response is about 0.85 on a scale from 0-1, indicating that respondents tend to agree or strongly agree with this statement.  As expected, women score more highly than men.  People answering the survey in languages other than French or Fulfulde are also slightly more supportive of women having a say in household finances than people answering questions in French or Fulfulde.

```{r wom_money}
wom_money.plot
```

### Women Work for Pay

The second question in the women's empowerment sequence asks if children suffer when women work for pay.  The responses have been re-scaled in the analysis so that a 1 means a pro-empowerment response (strongly disagree) and a 0 means an anti-empowerment response (strongly agree).  The mean score is about 0.73 on a scale from 0-1, indicating support for women, but less support than for the previous women's empowerment question.  However, this could easily be due to the reverse scaling of this question (making agreement normatively bad), and a persistent "agreement bias" wherein respondents are more likely to agree than disagree with survey statements.   Again, women score higher than men.  This time the non-French/Fulfulde speakers score lower than French/Fulfulde speakers^[Due to the relatively low number of non-French/Fulfulde speakers, there is much more random noise in the measurement of their opinions.]

```{r wom_work}
wom_work.plot
```

### Education of Boys and Girls

The third question in the women's empowerment sequence asks respondents if education is more important for a boy than a girl.  Unlike previous questions, there are _no statistical differences_ between men and women.  Women score slightly higher, but differences this small occur frequently due to chance.  Differences appear by region and language spoken: people in the northern region and people answering the survey in Fulfulde are more likely to agree that education is more important for boys than girls.

```{r eduboys}
eduboys.plot
#eduboy.all.lm<-svylm.fun(outcome='form.Attitudes.boy_edu_better',treatment='gender+adult+region+religion+form.End.survey_language',data=svy)
#eduboy.gen.lm<-svylm.fun(outcome='form.Attitudes.boy_edu_better',treatment='gender',data=svy) # no gender diff unless control for other things.
```

### Early/Child Marriage

<!--Chris: come back to this.-->
The fourth and final question in our women's empowerment index asks respondents if "It is _not_ in the best interest of a girl to be married before she is 16 years old."  No meaningful statistical differences.  Other religion significantly worse, but few people from that religion.

```{r earlymarriage}
earlymarriage.plot
#girlmarry.all.lm<-svylm.fun(outcome='form.Attitudes.boy_edu_better',treatment='gender*adult+form.End.survey_language',data=svy)
```

## Religious Tolerance

We ask four questions to measure beliefs about an individual's religion and that individual's perceptions of other religions.  Taken together, these four questions measure <!--"religious tolerance" -- something something...Chris/Rebecca: come up with a better summary name-->.  The individual questions are (1) There is more than one valid interpretation of religion, (2) My religion promotes peace and tolerance for members of other religions, (3) Other religions promote peace and tolerance for members of my religion, and (4) People of different religions live together peacefully in my community.  For each question, higher scores are normatively desirable and indicate more religious tolerance.  These four questions form an internally cohesive index with an alpha of 0.73.

All subsets of respondents score very highly on these measures of religious tolerance.  The average score on a 0-1 scale is `r round(mean(cpdf$reltol.index,na.rm=T),2)`, and all subsets analyzed score similarly high.  Overall there appears to be widespread religious tolerance in the north and extreme north regions of Cameroon.

Though these questions for a reasonably strong index, these questions measure distinct outcomes in themselves.  

```{r tol}
reltol.plot

#summary(cpdf$reltol.index)/max(cpdf$reltol.index,na.rm=T)
#reltol.all.lm<-svylm.fun(outcome='reltol.index',treatment='gender+adult+region+religion+form.End.survey_language',data=svy)
valid.plot
myrel_peace.plot
otherrel_peace.plot
diffrel_peace.plot
```

## Political/Civic Engagement

Does not form a reliable index.

```{r civic}
pol_trans.plot
dont_vote.plot
corruption.plot
comm_solve.plot

```

## Rando exp 

No plot.  Just didn't work.  Virtually everyone said yes in all conditions.

## Rel/Ethnic Feelings

Write stuff here.  These also really didn't work, since almost everyone says they overcome problems and don't vote based on ethnicity.

```{r rel/ethnic feelings}
vote_ethnic.plot
ethnic_divide.plot
religious_divide.plot
```

## Anti-Violence Empowerment

```{r anti-vio empow}
pos_change.plot
violence_problem.plot
violence_problem.plot
```

## Justice System

```{r justice system}
legal_recourse.plot
```

## Youth-Old

```{r youth-old}
youth.plot
old_respect.plot
youth_respect.plot
elders_understand.plot
elders_applicable.plot
youth_involved.plot
```



## Violence

Higher scores good, lower scores bad.

```{r vio}
vio.plot
defend_rel.plot
maintain.culture.plot
criminals_justice.plot
gov_change.plot
```


# Media Access

## Utilities

```{r electricity}
electricity.plot
```

## Phones/Tech Access

### Mobile phones

Youth more than adults.  North more than Ex north.  Muslim more than Chrstian.  Fulfulde more than French, both _way_ more than other language.

```{r mobile}
mobile.plot
```

### Smartphones

French way more, men way more, North way more.  Christian more. Adults more.

```{r smartph}
smartphone.plot
```

### Computers

Only 53 people have access to a computer.  Males more likely than women to have access.  French speaking more likely than other languages.

```{r comp}
computer.plot
#tally(cpdf$computer) # 53 
```

### Tablets

Only 14 people have tablets, so these numbers mean little.  Men more than women.  French more than other languages.

```{r tab}
tablet.plot
#tally(cpdf$tablet)
```

## Mobile Service Providers

```{r msp}
mtn.plot
camtel.plot
orange.plot
nextel.plot
otherMSP.plot
```

## OS

```{r OS}
android.plot
blackberry.plot
ios.plot
windows.plot
```

## Internet Use

Zero people used the internet on tablets.

```{r internet}
int.plot
int_friend_rel.plot
int_cafe.plot
int_mobile.plot
#int_tablet.plot # no one uses this
int_other.plot
```

## Media/Radio Listening

Only 9 people listened to Dabalaye.  Only 3 people listened to Chabab.  0 people listened to Dandal_Kura.

```{r radio programs}
don_derkeen.plot
dd_freq.plot
dd_op.plot
```

## Arewa24 Plots

```{r arewa}
alawar.plot
waiwaye.plot
dadin.plot
hhh.plot
kundin.plot
tauraruwa.plot
gari.plot
matasa.plot
other.plot
```


# Conclusion
 
Sum up findings here.

# Super Caveat

Mass Agreement Bias.  Almost everyone just agrees or says yes to every question.  This is a call for reverse coding and maybe also for more professional enumerators.  The data quality appears quite low.  OR the radio programs of PDEV really worked and these people are all super tolerant and stuff.  Well, they might be tolerant.  Or they might just know how to respond to surveys like this.
